MSF_Groupwork-NBA_shotlog

### 2016-10-10 MSFGroupwork-NBAHotHand.Rmd
PART 3 - HOT HAND THEORY

#select data with only scored result
#Point type (2 or 3 points) and distance
point_type <- c()
distance <- c()
shot_dist <- c()
for (z in 1:nrow(df)){
  if(df[z, "SHOT_RESULT"] == 1 & df[z, "PTS_TYPE"] >= 1){
    point_type <- c(point_type, df[z, "PTS_TYPE"])
    distance <- c(distance, df[z, "CLOSE_DEF_DIST"])
    shot_dist <- c(shot_dist, df[z, "SHOT_DIST"])
  }
}

#combine point type and distance
valid_distance = data.frame(point_type, distance, shot_dist)
dist_group <- c(5, 10, 15, 20, 25, Inf)
valid_distance[, "Range Group"] <- c(findInterval(valid_distance$distance, dist_group))

@@

#calculate correlation between defender distance and shot distance (2 or 3 points)
cor(distance, shot_dist)

@@

ggplot(data = valid_distance, aes(x = distance, y = shot_dist)) + geom_point()

ggplot(data = valid_distance, aes(x = distance) + geom_bar())

@@

#combine 2 point players and defender distance
player2point <- data.frame(players2, distance2)

summary(player2point)

#calculate mean distance from 2-pointers
avg_dist2 <- aggregate(distance2 ~ players2, player2point, mean)

@@

player_and_game <- function(x, y) {
   player <- df[df[ ,'player_id'] == x & df[ ,'GAME_ID1'] == y, 'FGM']
   number_of_streaks(as.vector(player))
}
player_and_game(202681, 21400681)

ourtable <- head(pttsum2,10)

list_per_player <- list()
j <- 1
for(i in ourtable$player_id) {
   streak_per_player[[j]] <- player_and_game(i, ourtable[ourtable[ ,'player_id'] == i, 'GAME_ID1'])
   j <- j + 1
}
streak_per_player

>>>>

#select data with only scored result
#Point type (2 or 3 points) and distance
point_type <- c()
distance <- c()
for (z in 1:nrow(df)){
  if(df[z, "SHOT_RESULT"] == 1 & df[z, "PTS_TYPE"] >= 1){
    point_type <- c(point_type, df[z, "PTS_TYPE"])
    distance <- c(distance, df[z, "CLOSE_DEF_DIST"])
  }
}

#combine point type and distance
valid_distance = data.frame(point_type, distance)

@@

@@

@@

#combine 2 point players and defender distance
player2point <- data.frame(players2, distance2)
summary(player2point)

@@

#Create a function that gives us a vector of streaks for a specific player and game
player_and_game <- function(x, y) {
   player <- df[df[ ,'player_id'] == x & df[ ,'GAME_ID1'] == y, 'FGM']
   number_of_streaks(as.vector(player))
}

#Set a new data.frame for the 10 players with best performances(higher points)
ourtable <- head(pttsum2,10)

#Create a list that gives us the streaks of each player and game (10 best performances)
streak_per_player <- list()
j <- 1
for(i in ourtable$player_id) {
   streak_per_player[[j]] <- player_and_game(i, ourtable[ourtable[ ,'player_id'] == i, 'GAME_ID1'])
   j <- j + 1
}
streak_per_player

#Create a data frame for the strikes of each one of the 10 players (cannot create a data.frame including all of them, because each list has different length)
player1_df <- data.frame('streaks' = streak_per_player[[1]])
player2_df <- data.frame('streaks' = streak_per_player[[2]])
player3_df <- data.frame('streaks' = streak_per_player[[3]])
player4_df <- data.frame('streaks' = streak_per_player[[4]])
player5_df <- data.frame('streaks' = streak_per_player[[5]])
player6_df <- data.frame('streaks' = streak_per_player[[6]])
player7_df <- data.frame('streaks' = streak_per_player[[7]])
player8_df <- data.frame('streaks' = streak_per_player[[8]])
player9_df <- data.frame('streaks' = streak_per_player[[9]])
player10_df <- data.frame('streaks' = streak_per_player[[10]])
ourtable

#Visualisation for 10 players - distribution of each players' streaklengths
ggplot(data = player1_df, aes(x = streaks)) + geom_histogram(binwidth = 1) + labs(title = ourtable[1,'player_name'])
ggplot(data = player2_df, aes(x = streaks)) + geom_histogram(binwidth = 1) + labs(title = ourtable[2,'player_name'])
ggplot(data = player3_df, aes(x = streaks)) + geom_histogram(binwidth = 1) + labs(title = ourtable[3,'player_name'])
ggplot(data = player4_df, aes(x = streaks)) + geom_histogram(binwidth = 1) + labs(title = ourtable[4,'player_name'])
ggplot(data = player5_df, aes(x = streaks)) + geom_histogram(binwidth = 1) + labs(title = ourtable[5,'player_name'])
ggplot(data = player6_df, aes(x = streaks)) + geom_histogram(binwidth = 1) + labs(title = ourtable[6,'player_name'])
ggplot(data = player7_df, aes(x = streaks)) + geom_histogram(binwidth = 1) + labs(title = ourtable[7,'player_name'])
ggplot(data = player8_df, aes(x = streaks)) + geom_histogram(binwidth = 1) + labs(title = ourtable[8,'player_name'])
ggplot(data = player9_df, aes(x = streaks)) + geom_histogram(binwidth = 1) + labs(title = ourtable[9,'player_name'])
ggplot(data = player10_df, aes(x = streaks)) + geom_histogram(binwidth = 1) + labs(title = ourtable[10,'player_name'])

#We can see that the distribution of all players is right skewed and that all of them (especially player3) have some long shooting streaks. To prove/disprove the hot hand theory we need to examine the indipendence of the shots. If each shot that a player takes is independent of the next shot, then a missed/made shot will not affect the propability of succesfulness of the next shot. So, if consecutive shots are independent, we disprove hot hand theory, while if consecutive shots are dependent, we prove the hot hand theory.
#Simulation
shot_results <- c('made', 'missed')
simul_shooter <- sample(shot_results, size = nrow(player1_df), replace = TRUE, prob = c(0.45, 0.55))
simul_shooter
```
